More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-11-14 2016-Weak-FH-Bisim/WeakBisim.tex
Proof of lemma 5

       Consider each premise of the open transition (constructed by \TrDeux\ in 
Definition~\ref{def:operationalSemantics}). 
We know each premise is \True for $P$ and try to prove the equivalent premise for 
$P[Q]_{j_0}$ (using the open transition of $Q$). 
$K$ and the synchronisation vector are unchanged ($j_0$ is now in the set of sub-pNets); 
$SV_k=\SV{(\alpha'_j)^{j\in I\uplus\{j_0\}\uplus 
	J}}{\alpha'}{e_b}$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. $I$ and $J$ 
	are the 
    set of leaves and holes of $P$, $I_1\uplus I_2$ and $J'$ are the sets of moving 
    leaves and holes 
    in the reduction of $P$. All sub-pNets of 
    must
    be 
reduced, we need:\\[-2ex]%\footnote{again we skip arguments related to renaming of fresh 
%variables}:
\[
    	\forall m\!\!\in\!\! I_1\uplus\{j_0\}. {\pNet_m 
    	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \]\[
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\]
%\forall m\!\!\in\!\! I\cup \{j_0\}. 	
%    {\left(\begin{array}{l}
%	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
%    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
%    	\{\xrightarrow{\beta_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
%    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
%    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
%		{ \pNet_m 
%    	 \models
%    	\inferrule*[myfraction=\reddottedrule]
%    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
%    	{\ostate{s_m} \OTarrow {v_m}
%    		\ostate{s_m'}} \land I_m=\{m\} \land J_m=\emptyset}
%\end{array}\right)}\]
the sub-pNet at position $j_0$ is the one filled by $Q$ which offers an open transition 
by hypothesis, the other open transitions are immediate consequence of the open 
transition that can be performed by $P$ (premises of \TrDeux).
The set of moving leaves is the union of the moving leaves in the open transition for $P$ 
and the ones for $Q$: 
$I'_{PQ}= I'_P \uplus I'_Q$, similarly the moving holes are the union of the moving 
holes, minus $j_0$: $J'_{PQ}=  J\setminus\{j_0\}\uplus J_Q$. The 
predicate for the open 
transition is: $\Pred\,''=\bigwedge_{m\in I _1\uplus I_2}\Pred_m \land \Pred\,'
\land \Pred(SV_k,v_i^{i\in I_1\uplus I_2}\uplus(j_0\mapsto v_Q),\beta_j^{j\in 
J},\alpha)$. 
By definition 
$\Pred(SV_k,\alpha_i^{i\in I}\uplus(j_0\mapsto \alpha_Q),\beta_j^{j\in 
J},v)\Leftrightarrow
	\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\alpha'_j \land 
	\alpha=\alpha'\land \alpha_Q=\alpha_{j_0}\land e_b$,
this is equivalent to $\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, 
\beta_j=\alpha'_j \land 
	\alpha=\alpha'\land \alpha_Q=\beta_{j_0}\land \beta_{j_0}=\alpha_{j_0}\land e_b$
 and by definition of $\Pred$ (as obtained by \TrDeux),
	$\Pred\,''\iff \Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$. The post-condition 
	gathers the post-conditions related to all 
	the leaves. Finally, the composed open transition can be
        built. \qed
        \end{proof}
\end{small}

@@

\subsection{Composition properties}
\TODO{Ludo: in the next 2 theorems I have a doubt about the biguplus of post p : not sure they should be disjoint, but else we need to do a complex composition  of posts}
\begin{lemma}[Weak OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction. Suppose $\gamma_{j_0}=[\beta_0..\beta_n]$
		 then there exist $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{\begin{array}{l}
p\in[0..n]
\end{array}} J_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p= \beta_p\land \Pred_p) \qquad\text{and~}
\Post=\Post\,'\uplus\!\! \biguplus_{p=0}^n
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}
\LUDO{A V2 by Ludo} \TODO{We will check notation  $\alpha_p=...$}

\begin{proof}
By lemma \ref{lem-rel-OT-WOT} we have: 
\begin{mathpar}
 P[Q]_{j_0} \models \openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}} 
\end{mathpar}
implies 	\\			
$			
	P[Q]_{j_0}	\models		
\Big(\forall i \in [0\ldots n].\openrule
    {
       \set{\gamma^1_i},\Pred^1_i,\Post^1_i   }
         {\ostate{s^1_i} \OTarrow {\tau} \ostate{s^1_{i+1}}}\, \wedge\,
\openrule
         {
           \set{\gamma^2},\Pred\,^2,\Post^2 }
         {\ostate{s^2} \OTarrow {\alpha} \ostate{s'^2}}
\,\wedge\,
\forall i \in [0\ldots m].\openrule
         {
           \set{\gamma^3_i},\Pred\,^3_i, \Post^3   }
         {\ostate{s^3_i} \OTarrow {\tau} \ostate{s^3_{i+1}}}
\Big) 
$

such that...\\

>>>>

       Consider each premise of the open transition (constructed by \TrDeux\ in 
Definition~\ref{def:operationalSemantics}). Let $P=\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle$
	\begin{description}
		\item[{\TrDeux:}]
	\end{description}

	\noindent
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
\end{mathpar}

We know each premise is \True\ for $P$ and try to prove the equivalent premise for 
$P[Q]_{j_0}$ (using the open transition of $Q$). 
$K$ and the synchronisation vector are unchanged ($j_0$ is now in the set of sub-pNets); 
$SV_k=\SV{(\alpha'_j)^{j\in I\uplus\{j_0\}\uplus 
	J}}{\alpha'}{e_b}$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. $I$ and $J$ 
	are the 
    set of leaves and holes of $P$, $I_1\uplus I_2$ and $J'$ are the sets of moving 
    leaves and holes 
    in the reduction of $P$. All sub-pNets of 
    must
    be 
reduced, we need:\\[-2ex]%\footnote{again we skip arguments related to renaming of fresh 
%variables}:
\[
    	\forall m\!\!\in\!\! I_1\uplus\{j_0\}. {\pNet_m 
    	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \]\[
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\]
%\forall m\!\!\in\!\! I\cup \{j_0\}. 	
%    {\left(\begin{array}{l}
%	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
%    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
%    	\{\xrightarrow{\beta_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
%    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
%    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
%		{ \pNet_m 
%    	 \models
%    	\inferrule*[myfraction=\reddottedrule]
%    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
%    	{\ostate{s_m} \OTarrow {v_m}
%    		\ostate{s_m'}} \land I_m=\{m\} \land J_m=\emptyset}
%\end{array}\right)}\]
the sub-pNet at position $j_0$ is the one filled by $Q$ (we define $P_{j_0}=Q$ and similarly $J_m=J_Q$, $\Pred_m=\Pred\,'$, $\Post_m=\Post\,'$,\ldots are the elements of the OT of Q) which offers an open transition 
by hypothesis, the other open transitions are immediate consequence of the open 
transition that can be performed by $P$ (premises of \TrDeux).
The set of moving leaves is the union of the moving leaves in the open transition for $P$ 
and the ones for $Q$; similarly the moving holes are the union of the moving 
holes, minus $j_0$: $J'_{PQ}=  J\setminus\{j_0\}\uplus J_Q$. The 
predicate for the open 
transition is: $\Pred\,''=\bigwedge_{m\in I _1\uplus I_2}\Pred_m \land \Pred\,'
\land \Pred(SV_k,v_i^{i\in I_1\uplus I_2}\uplus(j_0\mapsto v_Q),\beta_j^{j\in 
J},\alpha)$. 
By definition 
$\Pred(SV_k,\alpha_i^{i\in I}\uplus(j_0\mapsto \alpha_Q),\beta_j^{j\in 
J},v)\Leftrightarrow
	\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\alpha'_j \land 
	\alpha=\alpha'\land \alpha_Q=\alpha_{j_0}\land e_b$,
this is equivalent to $\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, 
\beta_j=\alpha'_j \land 
	\alpha=\alpha'\land \alpha_Q=\beta_{j_0}\land \beta_{j_0}=\alpha_{j_0}\land e_b$
 and by definition of $\Pred$ (as obtained by \TrDeux),
	$\Pred\,''\iff \Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$. The post-condition 
	gathers the post-conditions related to all 
	the leaves: $\biguplus_{m\in I_1\cup\{j_0\}\uplus I_2} 
    		\Post_m = 	\Post\uplus \Post\,'$. Finally, the composed open transition can be
        built by \TrDeux:

	\noindent
\begin{mathpar}
    \mprset {vskip=.8ex}
\inferrule
    {
\Leaves(\mylangle {\pNet}_m^{m\in I\cup\{j_0\}}, \set{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
\\
\\     	
	\forall m\!\!\in\!\! I_1\cup\{j_0\}. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J'_{PQ}=  J\setminus\{j_0\}\uplus J_Q	\\
    	\Pred\,'' =\bigwedge_{m\in I _1\uplus I_2}\Pred_m \land \Pred\,'
\land \Pred(SV_k,v_i^{i\in I_1\uplus I_2}\uplus(j_0\mapsto v_Q),\beta_j^{j\in 
J},\alpha)\\ 
    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1\cup\{j_0\}}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in  J'_{PQ}}, \Pred\,'',  	\Post\uplus \Post\,'}
    		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L\uplus L_Q}}}
    	}
    }
\end{mathpar}  
 \qed
        \end{proof}
\end{small}

@@

\subsection{Composition properties}
\TODO{Ludo: in the next 2 theorems I have a doubt about the biguplus of post p : not sure they should be disjoint, but else we need to do a complex composition  of posts}
\begin{lemma}[Weak OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction. Suppose $\gamma_{j_0}=[\beta_0..\beta_n]$
		 then there exist $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p= \beta_p\land \Pred_p) \qquad\text{and~}
\Post=\Post\,'\uplus\!\! \biguplus_{p=0}^n
		\Post_p  
		\end{mathpar}
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}
\LUDO{A V2 by Ludo}

\begin{proof} Suppose we have: 
\begin{mathpar}
 P[Q]_{j_0} \models \openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}} 
\end{mathpar}
By lemma \ref{lem-rel-OT-WOT} this implies we have: \\			
$ \forall i \in [0\ldots n]	\,		
	P[Q]_{j_0}	\models		
\openrule
    {
       \set{\gamma^1_i},\Pred^1_i,\Post^1_i   }
         {\ostate{s^1_i} \OTarrow {\tau} \ostate{s^1_{i+1}}}$
~and~ $\forall i \in [0\ldots m] \, P[Q]_{j_0}	\models \openrule
         {
           \set{\gamma^3_i},\Pred\,^3_i, \Post_i^3   }
         {\ostate{s^3_i} \OTarrow {\tau} \ostate{s^3_{i+1}}}
$\\
and $ P[Q]_{j_0}	\models \openrule
         {
           \set{\gamma^2},\Pred\,^2,\Post^2 }
         {\ostate{s^2} \OTarrow {\alpha} \ostate{s'^2}}
$ where\\
$
s=s^1_0 \wedge s^1_n=s^2 \wedge s'^2 = s^3_0 \wedge s^3_m=s',
\displaystyle{\gamma_j^{j\in J_1}=\mybigdotcup_{i=0}^n \vis{\set{\beta^1_i}} \dotcup  \vis{\set{\beta^2}} \dotcup \mybigdotcup_{i=0}^m \vis{\set{\beta^3_i}}}
\\
\Pred_{OT}=\bigwedge_{i=0}^n\Pred^1_i\land\Pred\,^2\bigotimes_{i=0}^{n}\subst{\Post_i^1}\land (\bigwedge_{i=0}^m\Pred_i^3\subst{\Post^2})\bigotimes_{i=0}^{n}\subst{\Post^1_i}\\
\Post_{OT}=\biguplus_{i=0}^{m}(\Post_i^3\subst{\Post^2})\bigotimes_{i=0}^{n}\subst{\Post^1_i}
$
